Disk Heating, Galactoseismology, and the Formation of Stellar Halos

### 2017-06-27 paper/paper.tex
fix f RR MG

Since we were unable to find any RR Lyrae clearly associated with TriAnd 1, our attempt to measure a more accurate distance to this structure was unsuccessful.
However, the upper limit on the value of $\frrmg$ was in itself interesting.
RR Lyrae and M giants are tracers of populations with quite distinct metallicities: stars in the horizontal branch phase of evolution are typically only blue enough to cross the instability strip and become RR Lyrae if they have [Fe/H]$\lesssim -1.5$; and giant stars typically only evolve to colors red enough to become spectral class M if they have [Fe/H]$\gtrsim$-1.5.
Hence, a stellar population has to contain a significant range of metallicities in order to contain both types of stars.
%In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
%It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd 1 and 2 had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \cite[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \cite{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \cite[i.e. $f_{\rm RR:M} \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

@@

The origin of stellar associations can also be explored through the chemical compositions of their constituent stars.
However, our finding (described in Section \ref{sec:populations}) that the low-latitude-structures all have stellar populations (as indicated by $f_{\rm RR:M}$) that look more like the disk than known Galactic satellites is at odds with prior work on abundance patterns.
For example, high-resolution spectra of 21 M giants in Mon/GASS \cite{chou2010b} showed
[Ti/Fe] lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk \cite[e.g.,][]{reddy03,bensby2014} and a mean offset for [Y/Fe] of about 0.2 dex at ${\rm [Fe/H]} \approx -0.5$. A comparison with results from the same group for stars in the Sgr dwarf spheroidal galaxy \cite{chou2010a} suggested more similar patterns to those seen in Mon/GASS and support  an external origin for this structure. Subsequent work comes to similar conclusions for the TriAnd Clouds \cite{chou11}.

@@

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low latitude structures --- Mon/GASS, TriAnd 1/2 and A13  --- each have low velocity dispersions supporting the genuine association of the stars within them;
\item Mon/GASS, TriAnd 1/2 and A13 also share a single, continuous sequence in average velocity as a function of $l$ suggestive of associations across these groups;
\item the stellar populations (as indicated by $f_{\rm RR:M}$) in these groups look like each other;
% ALLY COMMENT: Combine with next point? “look like each other” is a bit vague!
\item the stellar populations are more consistent with those in the Galactic disk rather than those observed in the stellar halo or Galactic  satellites;
\item the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy;
\item the low-latitude structures around the outer Galaxy can be connected to overdensities oscillating between the northern and southern hemispheres on smaller scales and traced all the way back to the Solar Neighborhood;
\item small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood which suggest the local disk is moving up and down.
\end{itemize}

>>>>

Since we were unable to find any RR Lyrae clearly associated with TriAnd 1, our attempt to measure a more accurate distance to this structure was unsuccessful.
However, the upper limit on the value of $\frrmg$ was in itself interesting.
RR Lyrae and M giants are tracers of populations with quite distinct metallicities: stars in the horizontal branch phase of evolution are typically only blue enough to cross the instability strip and become RR Lyrae if they have [Fe/H]$\lesssim -1.5$; and giant stars typically only evolve to colors red enough to become spectral class M if they have [Fe/H]$\gtrsim$-1.5.
Hence, a stellar population has to contain a significant range of metallicities in order to contain both types of stars.
%In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
%It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd 1 and 2 had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \cite[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \cite{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \cite[i.e. $\frrmg \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

@@

The origin of stellar associations can also be explored through the chemical compositions of their constituent stars.
However, our finding (described in Section \ref{sec:populations}) that the low-latitude-structures all have stellar populations (as indicated by \frrmg) that look more like the disk than known Galactic satellites is at odds with prior work on abundance patterns.
For example, high-resolution spectra of 21 M giants in Mon/GASS \cite{chou2010b} showed
[Ti/Fe] lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk \cite[e.g.,][]{reddy03,bensby2014} and a mean offset for [Y/Fe] of about 0.2 dex at ${\rm [Fe/H]} \approx -0.5$. A comparison with results from the same group for stars in the Sgr dwarf spheroidal galaxy \cite{chou2010a} suggested more similar patterns to those seen in Mon/GASS and support  an external origin for this structure. Subsequent work comes to similar conclusions for the TriAnd Clouds \cite{chou11}.

@@

The observations summarized in Section \ref{sec:obs} indicate that:
\begin{itemize}
\item the low latitude structures --- Mon/GASS, TriAnd 1/2 and A13  --- each have low velocity dispersions supporting the genuine association of the stars within them;
\item Mon/GASS, TriAnd 1/2 and A13 also share a single, continuous sequence in average velocity as a function of $l$ suggestive of associations across these groups;
\item the stellar populations (as indicated by \frrmg) in these groups look like each other;
% ALLY COMMENT: Combine with next point? “look like each other” is a bit vague!
\item the stellar populations are more consistent with those in the Galactic disk rather than those observed in the stellar halo or Galactic  satellites;
\item the abundance patterns  of the low-latitude structures are similar to those of the thick disk of our Galaxy;
\item the low-latitude structures around the outer Galaxy can be connected to overdensities oscillating between the northern and southern hemispheres on smaller scales and traced all the way back to the Solar Neighborhood;
\item small-scale, systematic vertical motions of and within the disk have been detected in the Solar Neighborhood which suggest the local disk is moving up and down.
\end{itemize}
